More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-10-04 2016-Weak-FH-Bisim/WeakBisim.tex
Remplacement Predsv + nettoyage des RAB
minor change
Back to rabea's merge
minor change again
notations

Let $\uplus$ be the disjoint union operator on indexed sets. 
The elements
of the union are then accessed by using an index of one of the two
joined families.
We suppose here that disjoint unions are always well defined; this
requires to choose disjoint  
indices in the definition of the component system (e.g. name of
methods of different interfaces).

$\subst{y\gets x}$ is the substitution operation that replaces in a term all occurrences 
of the term $y$ by the term $x$ (note that the absence of binders makes this operation 
trivial).

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. $v$ is a variable denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\subst{x_k\gets e_k}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

\RAB{diff entre Leaves(p) et overline\{Leaves\}?}

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{Sort},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Consider a synchronisation vector $SV_k$, for $k\in K$. We define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array}\]

@@

\RAB{ce pred est-il diff de celui de la def 4?}

@@

\begin{definition}[Operational semantics of open pNets]
	\label{def:operationalSemantics}
	The semantics of a pNet $p$ is an open automaton $A = <Leaves(p),J,\mathcal{S}, s_0,
	\mathcal{T}>$ where:
	\begin{itemize}
		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
		\item $\mathcal{T}$ is the smallest set of open transitions
		satisfying the rules below:
	\end{itemize}

@@

\RAB{notation S, T or underline\{S\}, underline\{T\}? pas clair...}

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\ SV\!\!= \!clone(SV_k) \!=\! \alpha_m^{m \in I_k\uplus J_k} \!\to\! 
    \alpha'_k \\
    	Leaves(p) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\in I_k.
    	\pNet_m \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \xrightarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}
    	%\land
    	%Leaves(\pNet_m) = \overline{pLTS}^{L_k})
    	\\
    	I' = \biguplus_{m\in I_k}\!\! I_m'
    	\\ J' = \biguplus_{m\in I_k}\!\! J'_m \uplus J_k 	\\
    	\Pred = \bigwedge_{m\in I_k}\!\! \Pred_m \land
    	\Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k},v)\\
    	\forall j\!\!\in\!\! J_k. {\tt 
    		fresh}(b_j) \\ {\tt fresh}(v) \\ 
    	\forall i\in
    	L\backslash I'.\,s'_i=s_i 
    }
    {\mylangle \pNet_i^{i\in I}, Sort_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \xrightarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $

\RAB{que signifie notation uplus Post? Pourquoi fresh(v), et v existe satisfait Pred? aussi b\_j?}

@@

@@

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea. Formally:

 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim}

\noindent
\begin{minipage}{0.69\linewidth} 	Suppose that
   $A_1 = <L_1,J,\mathcal{S}_1, s^1_0,
   \mathcal{T}_1>$ and $A_2 = <L_2,J,\mathcal{S}_2,s^2_0, \mathcal{T}_2>$
   are open automata where the set
   of holes are equal %\LUDO{added:} 
   and of the same sort.  %\LUDO{changed:} 
      Let 
      \ERIC{$\mathcal{R}=\{(s^1,s^2|\Pred_{s^1,s^2})\}$} be a
      relation over the sets 
      $\mathcal{S}_1$ and $\mathcal{S}_2$ 
      constrained by a predicate.  More precisely, for any pair $(s^1,s^2)$, there is a 
      single
      $(s^1,s^2|\Pred_{s^1,s^2})\in\mathcal{R}$  stating that $s^1$ and $s^2$ are related if $\Pred_{s^1,s^2}$\ 
      is 
      true.
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}
 Then 
   $\mathcal{R}$ is an FH-bisimulation iff for any  states
   $s^1\in\mathcal{S}_1$ and\\ $s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have
   the following:

@@

       	Let $\mathcal{R}''$ be the relation: 
       	\[\mathcal{R}'' = \{(s^1,s^3|\Pred\,'')|\,\Pred\,''=\bigvee_{ \begin{array}{l}
       		(s^1,s^2|\Pred)\in\mathcal{R}\\ (s^2,s^3|\Pred\,')\in\mathcal{R}' 	
       		\end{array}
       	}\,\Pred\land\Pred\,'\}\]

This relation is the adaptation of the transitivity to the conditional relationship that 
defines a bisimulation. Indeed the global disjunction together with the conjuction of 
predicates plays exactly the role of the intermediate element in a transitivity rule: 
``there exists an intermediate state'' corresponds to the global disjunction, and the 
conjunction of states expresses the intermediate predicate is used to ensure 
satisfiability of the predicate relating the first state to the last one..

>>>>

Let $\uplus$ be the disjoint union operator on indexed sets. 
The elements
of the union are then accessed by using an index of one of the two
joined families.
We suppose here that disjoint unions are always well defined; this
requires to choose disjoint  
indices in the definition of the component system (e.g. name of
methods of different interfaces).
We denote $y\gets x$ a substitution. Application of the substitution is denoted
$\subst{y\gets x}$, the operation replaces in a term all occurrences 
of the term $y$ by the term $x$ (note that the absence of binders makes this operation 
trivial).

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. $v$ is a variable denoting the resulting action
        of this open transition. \Pred\ is a predicate 
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{Sort},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Consider a synchronisation vector $SV_k$, for $k\in K$. We define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred_{\symb{sv}}(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array}\]

@@

@@

\begin{definition}[Operational semantics of open pNets]
	\label{def:operationalSemantics}
	The semantics of a pNet $p$ is an open automaton $A = <Leaves(p),Holes(p),States(p),InitState(p),
	\mathcal{T}>$ where $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\begin{itemize}
%		\item $J$ is the set of holes: $Holes(p)= J$. 
		%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
%		\item ${\mathcal{S}} = States(p)$ and $s_0 = InitState(p)$
%		\item $\mathcal{T}$ is the smallest set of open transitions		satisfying the rules below:
%	\end{itemize}

@@

@@

	\noindent
    $\inferrule
    {k\!\in\! K \\ SV\!\!= \!{\tt clone}(SV_k) \!=\! \alpha_m^{m \in I_k\uplus J_k} \!\to\! 
    \alpha'_k \\
    	Leaves(p) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\in I_k.
    	\pNet_m \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \xrightarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}
    	%\land
    	%Leaves(\pNet_m) = \overline{pLTS}^{L_k})
    	\\
    	I' = \biguplus_{m\in I_k}\!\! I_m'
    	\\ J' = \biguplus_{m\in I_k}\!\! J'_m \uplus J_k 	\\
    	\Pred = \bigwedge_{m\in I_k}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV,a_i^{i\in I_k},b_j^{j\in J_k},v)\\
    	\forall j\!\!\in\!\! J_k. {\tt 
    		fresh}(b_j) \\ {\tt fresh}(v) \\ 
    	\forall i\in
    	L\backslash I'.\,s'_i=s_i 
    }
    {\mylangle \pNet_i^{i\in I}, Sort_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \xrightarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $

@@

\RAB{Exemple a revoir. Les regles ne passent...}

@@

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea.\\
\TODO{To check:}
Let \ERIC{$\mathcal{R}=\{(s^1,s^2|\Pred_{s^1,s^2})\}$} be a relation over the set $\mathcal{S}_1$ and $\mathcal{S}_2$ constrained by a predicate
More precisely, for any pair $(s^1,s^2)$, there is a 
   single
      $(s^1,s^2|\Pred_{s^1,s^2})\in\mathcal{R}$  stating that $s^1$ and $s^2$ are related if $\Pred_{s^1,s^2}$\ 
      is 
      true.

\RAB{ I propose:
Let $\mathcal{R}=\{(s^1,s^2|\Pred_{s^1,s^2})\}$ be a relation over the set $\mathcal{S}_1$ and $\mathcal{S}_2$ constrained by a predicate: if $(s^1,s^2) \in \mathcal{R}$ then $\Pred_{s^1,s^2}$ is  true.}

 FH-bisimulation is defined formally: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim}

\noindent
\begin{minipage}{0.69\linewidth} 	Suppose that
   $A_1 = <L_1,J,\mathcal{S}_1, s^1_0,
   \mathcal{T}_1>$ and $A_2 = <L_2,J,\mathcal{S}_2,s^2_0, \mathcal{T}_2>$
   are open automata where the set
   of holes are equal %\LUDO{added:} 
   and of the same sort.  %\LUDO{changed:} 
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}
 Then 
   $\mathcal{R}$ is an FH-bisimulation iff for any  states
   $s^1\in\mathcal{S}_1$ and\\ $s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have
   the following:

@@

       	Let $\mathcal{R}''$ be the relation: 
       	\[\mathcal{R}'' = \{(s^1,s^3|\Pred\,'')\,\,\Big|\,\Pred\,''=\bigvee_{\begin{array}{c}       		
       		(s^1,s^2|\Pred)\in\mathcal{R}\\ (s^2,s^3|\Pred\,')\in\mathcal{R}' 	
       		\end{array}
       	}\,\Pred\land\Pred\,'\}\]

This relation is the adaptation of the transitivity to the conditional relationship that 
defines a bisimulation. Indeed the global disjunction together with the conjunction of 
predicates plays exactly the role of the intermediate element in a transitivity rule: 
``there exists an intermediate state'' corresponds to the global disjunction, and the 
conjunction of states expresses the intermediate predicate is used to ensure 
satisfiability of the predicate relating the first state to the last one..
